---
title: "Description"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Description}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
# Load and install SwineNet package

```{r setup}
list.of.packages <- c("SwineNet")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

lapply(list.of.packages, require, character.only = TRUE)
```

Load example datasets

```{r}
data(sites)
data(movements)
```
Define a list of attributes for premises to keep in the network

```{r}
premises.attributs <- c("siteID", "farmGroup",  "type", "freerange", "BRS", "lon", "lat", "mainland")
```
Add year, month and semester to the movement data.frame using timewindows function.
```{r}
movements <- timewindows(movements, formatDate = "%Y-%m-%d")
```

# Build network

Produce igraph graph using the movements data.frame  (both first columns should respectivelly be named "from" and "to" and should be consistent with the premises ids) and the premises data.frame (site id has to be the first column). Arguments 'path.save' and 'prefixe' are requested only if save.network is TRUE. By default the duplicated edges will be kept and loops will be removed, use arguments: remove.duplicated.edges and remove.loop.edges to modify.

```{r }
<<<<<<< HEAD
G <- get_network(movements = movements,
=======
G <- generate_network(movements = movements,
>>>>>>> 42f1e3243b0b136dd72c76edb43e40e46dfee4bb
                      premises = sites[,premises.attributs],
                      save.network = FALSE,
                      remove.duplicated.edges = F,
                      remove.loop.edges = T)
```

## Limit analysis to a subset of premises

To limit the analysis to a subset of premises of edges. Update the premises and/or movements data.frames.
For example: select premises on the mainland
```{r }
sites <- sites[which(sites$mainland==TRUE),]
premises.attributs <- c("siteID", "farmGroup",  "type", "freerange", "BRS", "lon", "lat")
```

### Subset of farms: select active premises of the national industry in the mainland

```{r }
selectedFARMS <- sites[which(sites$type %in% c("FA", "FPW", "FF", "PW", "PWF", "FI", "MU", "NU","BS","SL")),premises.attributs]

related.movements <- movements[which(movements$from %in% selectedFARMS$siteID &
                                               movements$to %in% selectedFARMS$siteID ),]


<<<<<<< HEAD
G <- get_network(movements = related.movements,
=======
G <- generate_network(movements = related.movements,
>>>>>>> 42f1e3243b0b136dd72c76edb43e40e46dfee4bb
                      premises = selectedFARMS[,premises.attributs],
                      save.network = FALSE)
```

Repeat the operation excluding slaughterhouses and boar stations

```{r }
selectedFARMS <- sites[which(sites$type %in% c("FA", "FPW", "FF", "PW", "PWF", "FI", "MU", "NU")),
                              premises.attributs]

related.movements <- movements[which(movements$from %in% selectedFARMS$siteID &
                                           movements$to %in% selectedFARMS$siteID ),]

<<<<<<< HEAD
G <- get_network(movements = related.movements,
=======
G <- generate_network(movements = related.movements,
>>>>>>> 42f1e3243b0b136dd72c76edb43e40e46dfee4bb
                      premises = selectedFARMS,
                      save.network = FALSE)
```

## Generate subnetworks based on edge attribute (such as various time windows or type of transport)

<<<<<<< HEAD
Use splitByEdges term of the 'get_network' function. 'splitByEdges' attribute should refer to a qualitative variable in movements data.frame.
=======
Use splitByEdges term of the 'generate_network' function. 'splitByEdges' attribute should refer to a qualitative variable in movements data.frame.
>>>>>>> 42f1e3243b0b136dd72c76edb43e40e46dfee4bb
Note that, the general network with all movements can be include by adding a NA object in the splitByEdges vector.

```{r }
selectedFARMS <- sites[which(sites$type %in% c("FA", "FPW", "FF", "PW", "PWF", "FI", "MU", "NU")),
                              premises.attributs]

related.movements <- movements[which(movements$from %in% selectedFARMS$siteID &
                                               movements$to %in% selectedFARMS$siteID ),]

<<<<<<< HEAD
G <- get_network(movements = related.movements,
=======
G <- generate_network(movements = related.movements,
>>>>>>> 42f1e3243b0b136dd72c76edb43e40e46dfee4bb
                      premises = selectedFARMS,
                      splitByEdges = c("MType", "year"),
                      save.network = FALSE)

G %>% names
```
## Generate subnetworks based on vertex attribute (type of farms)

<<<<<<< HEAD
Use splitByVertex term of the 'get_network' function. 'splitByVertex' attribute should refer to a qualitative variable in premises data.frame.
=======
Use splitByVertex term of the 'generate_network' function. 'splitByVertex' attribute should refer to a qualitative variable in premises data.frame.
>>>>>>> 42f1e3243b0b136dd72c76edb43e40e46dfee4bb
Note that, the general network with all Premises can be include by adding a NA object in the splitByVertex vector.
Note that, splitByEdges and splitByVertex can be used jointly and networks will be produced for all combinations of movement subsets and premise subsets.

```{r }

<<<<<<< HEAD
G <- get_network(movements = related.movements,
=======
G <- generate_network(movements = related.movements,
>>>>>>> 42f1e3243b0b136dd72c76edb43e40e46dfee4bb
                      premises = selectedFARMS,
                      splitByVertex = "type",
                      save.network = FALSE)

G %>% names

<<<<<<< HEAD
G <- get_network(movements = related.movements,
=======
G <- generate_network(movements = related.movements,
>>>>>>> 42f1e3243b0b136dd72c76edb43e40e46dfee4bb
                      premises = selectedFARMS,
                      splitByEdges = c(NA, "semester"),
                      splitByVertex = "freerange",
                      save.network = FALSE)

G %>% names
```


# Network representations
Networks can be represented based on its attributes with vis_network function.

```{r eval=FALSE}
igraph::get.edge.attribute(G[[1]]) %>% names
igraph::get.vertex.attribute(G[[1]]) %>% names

p <- vis_network(networktoplot = G[[1]],
                        title = "network representation",
                        movements.type.lab = "MType",
                        premises.type.lab = "type")

p
```

If coordinates are included as lat and lon vertex attributes, networks can also be mapped using map_network function.
Use the map_network function to map the network.
```{r}
igraph::get.vertex.attribute(G[[1]]) %>% names

map_network(
  network = G[[1]],
  country = c("FR"),
  premisescolours = "type",
  title = "network map",
  topfeature = "point"
)

```

# Network centralities

A distance matrix of pairwise of premises can be generated using the GeoDistanceInMetresMatrix function.

```{r}
distMatrix <- GeoDistanceInMetresMatrix(selectedFARMS[,c("siteID","lon","lat")])
# or in km
distMatrix <- round(GeoDistanceInMetresMatrix(selectedFARMS[,c("siteID","lon","lat")]) / 1000,1)
```

<<<<<<< HEAD
centralities_network function can be use to calculate a predefine basic set of centralities of the network, including or not teh distance matrix.
Use ```do.call(rbind, lapply(G, centralities_network))``` to generate centralities data.frame of centralities for multiple networks.

```{r}
#  without the distance
centralities_network(G[[1]])

#  or with the distance
centralities_network(network = G[[1]], include.distance = TRUE, distance.matrix = distMatrix)
=======
network_centralities function can be use to calculate a predefine basic set of centralities of the network, including or not teh distance matrix.
Use ```do.call(rbind, lapply(G, network_centralities))``` to generate centralities data.frame of centralities for multiple networks.

```{r}
#  without the distance
network_centralities(G[[1]])

#  or with the distance
network_centralities(network = G[[1]], include.distance = TRUE, distance.matrix = distMatrix)
>>>>>>> 42f1e3243b0b136dd72c76edb43e40e46dfee4bb
```
```

# Exponential random graph model

## Design the experiment
ergm package works with network, transformation of igraph graph into network can be made using intergraph package.

### Which covariables to include in the bidirectional stepwise procedure?
Define a list of covariates to include in the bidirectional stepwise procedure.
The list can contain terms related to nodal and edges attributes as well as structural terms.
Edges attributes can only be included with a matrix involving a value for each pair of premises (eg. distance matrix).

Use expertise, summary function and mixingmatrix function to define the covariables. Summary and mixingmatrix functions work with network, not igraph.

Definition for implemented terms can be found within the ergm package help (?ergm.terms).

```{r}
g <- intergraph::asNetwork(G[[1]])

########### 
########### Nodal and edges attributes
########### 
library(network)

network::list.vertex.attributes(g)

farmGroup.levels <- network::get.vertex.attribute(g, "farmGroup") %>% unique %>% length
type.levels <- network::get.vertex.attribute(g, "type") %>% unique %>% length
BRS.levels <- network::get.vertex.attribute(g, "BRS") %>% unique %>% length
freerange.levels <- network::get.vertex.attribute(g, "freerange") %>% unique %>% length

elements <- c(
  "edgecov(distMatrix)",
  paste("nodefactor(\"farmGroup\", levels=c(",seq(farmGroup.levels),"))"),
  paste("nodeifactor(\"farmGroup\", levels=c(",seq(farmGroup.levels),"))"),
  paste("nodeofactor(\"farmGroup\", levels=c(",seq(farmGroup.levels),"))"),
  paste("nodefactor(\"type\", levels=c(",seq(type.levels),"))"),
  paste("nodeifactor(\"type\", levels=c(",seq(type.levels),"))"),
  paste("nodeofactor(\"type\", levels=c(",seq(type.levels),"))"),
  paste("nodefactor(\"BRS\", levels=c(",seq(BRS.levels),"))"),
  paste("nodeifactor(\"BRS\", levels=c(",seq(BRS.levels),"))"),
  paste("nodeofactor(\"BRS\", levels=c(",seq(BRS.levels),"))"),
  paste("nodefactor(\"freerange\", levels=c(",seq(freerange.levels),"))"),
  paste("nodeifactor(\"freerange\", levels=c(",seq(freerange.levels),"))"),
  paste("nodeofactor(\"freerange\", levels=c(",seq(freerange.levels),"))"),
  # interactions
  paste("nodematch(\"farmGroup\", diff=F)"),
  paste("nodematch(\"BRS\", diff=F)"),
  paste("nodematch(\"farmGroup\", diff=T, levels=c(",seq(farmGroup.levels),"))"),
  paste("nodemix(\"freerange\", levels2=c(",seq(freerange.levels^2),"))"),
  paste("nodemix(\"BRS\", levels2=c(",seq(BRS.levels)^2,"))"),
  paste("nodemix(\"type\", levels2=c(",seq(type.levels^2),"))")#,
  # paste("asymmetric(\"type\")"),
  # paste("asymmetric(\"type\", diff = T, levels=c(",seq(type.levels),"))")
  ) %>% as.list

elements %>% unlist %>% length

########### 
########### Structural characteristics
########### 

library(sna)
ostar.levels <- degree(g,gmode = "out") %>% unique
istar.levels <- degree(g,gmode = "in") %>% unique
triadcensus.levels <- c(0:15)[which(summary(g ~ triadcensus(0:15)) != 0)]
ITP.levels <- which(summary(g~ddsp(1:50, "ITP")) !=0)
OTP.levels <- which(summary(g~ddsp(1:50, "OTP")) !=0)
OSP.levels <- which(summary(g~ddsp(1:50, "OSP")) !=0)

structural.elements <- as.list(c(
  "edges","twopath", "idegree1.5", "odegree1.5", "transitive",
  paste("ostar(",seq(ostar.levels),")"),
  paste("istar(",seq(istar.levels),")"),
  paste("triadcensus(",triadcensus.levels,")"),
  paste("ddsp(",seq(OSP.levels),", type =\"OSP\")"),
  paste("ddsp(",seq(ITP.levels),", type =\"ITP\")"),
  paste("ddsp(",seq(OTP.levels),", type =\"OTP\")")
))

structural.elements %>% unlist %>% length
```

## Launch the stepwise procedure
Define the baseline formula and launch the stepwise procedure either "forward", "backward" or "bidirectional".
If the function crashes, reload the last stepwise.summary output saved and relaunch the procedure the covariable that produced the crash will be skip for this step.
As the function call a future.apply function allowing for parallelisation, use closeAllConnections() after a crash, make sure that all connections have been closed before relaunch.

```{r eval=FALSE}
########### 
########### Run the stepwise selection process
########### 

base.formula <- "g ~ edges"

########### 
########### Forward stepwise procedure
########### 

stepwise4ERGM(base.formula, # string
                           nbworkers =  10, # integer
                           elements = append(elements,structural.elements), # list
                           network.name = "mynetwork", # string
                           results.path = getwd(), # string
                           maxduration = 60*60*1, # integer in seconds
                           g = g,
                           mode = "forward", # string "forward" or "backward"  or "bidirectional"
                           verbose =T
)



########### 
########### Backward stepwise procedure
########### 


load(file = getwd() %>% 
  list.files(., full.names = TRUE) %>% 
  .[grep(paste0("stepwise.summary_mynetwork"),.)]%>% 
  extract(which.max(file.mtime(.))))

base.formula <- stepwise.summary[[stepwise.summary %>% length]]$formula %>%  substr(., gregexpr("\\g ~", .)[[1]][1] , nchar(.)-1)
baseline.aic <- stepwise.summary %>% .[[length(.)]] %>% .$min.aic

stepwise4ERGM(base.formula,
                           nbworkers = 10, #integer
                           network.name = "mynetwork", # string
                           results.path = getwd(),
                           maxduration = 60*60*2, # integer in seconds
                           g,
                           mode = "backward", # string "forward" or "backward"  or "bidirectional"
                           stepwise.summary = stepwise.summary,
                           baseline.aic = baseline.aic, 
                           verbose = T
)

######## 
######## Bidirectional stepwise procedure
######## 

stepwise4ERGM(base.formula,
                           nbworkers =  10, # integer
                           elements = elements, # list
                           network.name = "mynetwork", # string
                           results.path = results.path,
                           maxduration = 60*60*1, # integer in seconds
                           g = g,
                           mode = "bidirectional", # string "forward", "backward" or "bidirectional"
                           verbose =T
)

```

## Analyse results


```{r eval = FALSE}
# read the most recent output
load(file = list.files(., full.names = TRUE) %>% 
  .[grep(paste0("stepwise.summary_mynetwork"),.)]%>% 
  extract(which.max(file.mtime(.))))

selected.terms <- analyseStepwise(stepwise.summary)
```

# Analyse the selected model

Use the selectedERGM function to analyse the most parcimonious combinations of covariables.
The function will save output files in the specified path.
Runing time can be long, be patient.

```{r eval = FALSE}
selectedERGM(stepwise.summary,
             g,
             outputpath='',
             distMatrix = NA,
             nsimGof = 2000,
             nSimNet = 5,
             maxtimeERGM = 60*60*2,
             country.plot = c("FR"),
             nrepAIC = 3,
             verbose = T)
```


